---
title: "Wellcome Trust Project - Exploratory Data Report"
subtitle: "Version 04"
output:
  html_document:
    toc: true
    toc_depth: 3
    collapsed: true
    toc_float: true
---

```{r setup, include=FALSE}

def <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
  x <- def(x, options)
  ifelse(!is.null(options$suppress), 
         gsub(pattern = "```.*```|\\[.*\\]|<p>\\[.*\\]</p>", "", x), x)
})


knitr::opts_chunk$set(echo = F, warning = F, message = F)



library(tidyverse)
library(knitr)

# change path to work on NatCen laptop
path <- "\\\\homerfp01/FolderRedirection/AlexC/Documents/work/connect"

# load data
alspac_long_full <- read_rds(str_c(path, "/data/clean/alspac_long_v02.rds"))

# select cases that participated at age 7
alspac_long <- alspac_long_full %>% 
  filter(baseline_age_7y9m == 1) 



# load functions
list.files(str_c(path, "/scripts/functions/"), full.names = T) %>% 
  map(source)

```
 
# Baseline sample

As a baseline sample we will use participants to the KR wave (age 7) given that our outcome variables are collected from this age. This comprises of 8,235 cases out of a total of 15,645:


```{r}
alspac_long_full %>% 
  filter(age == 8) %>% 
  count(baseline_age_7y9m) %>% 
  mutate(prop = n/sum(n),
         baseline_age_7y9m = factor(baseline_age_7y9m, 
                                    labels = c("No", "Yes"))) %>% 
  setNames(c("Participated at age 7", "N", "Proportion")) %>% 
  kable(digits = 2)
```

 
 
# Univariate statistics 

## Outcomes 

```{r results='asis'}

dep_change <- alspac_long %>% 
  select(age, anxiety, depression, depression_smfq) %>% 
  group_by(age) %>% 
  summarise(Anxiety_prop = mean(anxiety, na.rm = T),
            Anxiety_miss = mean(is.na(anxiety)),
            Depression_prop = mean(depression, na.rm = T),
            Depression_miss = mean(is.na(depression)),
            DepressionSmfq_prop = mean(depression_smfq, na.rm = T),
            DepressionSmfq_miss = mean(is.na(depression_smfq)))


dep_change %>% 
    pivot_longer(!age,
               names_sep = "_",
               names_to = c( "Variable", ".value")) %>% 
  na.omit() %>% 
  arrange(Variable, age) %>% 
  mutate(Percentage = prop * 100,
         `Missing (%)` = miss * 100) %>% 
  select(-prop, -miss) %>% 
  rename(Age = age) %>% 
  knitr::kable(digit = 2)
```


```{r}
dep_change %>% 
  select(age, ends_with("prop")) %>% 
  gather(value = value, key = var, -age) %>% 
  mutate(var = str_remove(var, "_prop"),
         age = as.factor(age)) %>% 
  na.omit() %>% 
  ggplot(aes(age, value, color = var, group = var)) + 
  geom_line() +
  geom_point() +
  theme_bw() +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Age",
       y = "Percentage of cases",
       color = "")

```

## Key predictors {.tabset}

### Bullying

```{r results='asis', suppress = T}
bully_var <- c("bullying_victim_fct", 
               "bullying_often_bullied_fct",
               "bullying_frequency_bullied_fct")
test <- map(bully_var, function(x) tab_desc_tab(.data[[x]]))

for (i in seq_along(test)) print(test[[i]])

```

```{r suppress = T}
map(bully_var, function(x) fig_desc_long(.data[[x]]))
```

### Friends

```{r results='asis', suppress = T}
friends_var <- c("friends_plays_with_kids_fct",
                 "friends_freq_visits_fct",
                 "friends_happy_num_friends_fct",
                 "friends_understand_fct",
                 "friends_overall_happy_fct",
                 "friends_talk_problems_fct",
                 "friends_freq_see_outside_school_fct",
                 "friends_whether_any_close_friend",
                 "friends_num_close_friend_fct",
                 "friends_yps_friends_fall_out_with_them_fct",
                 "friends_friends_support_yp_when_they_need_them_fct",
                 "friends_friends_put_yp_down_in_front_of_others_fct",
                 "friends_friends_make_yp_feel_confident_fct",
                 "friends_ch_one_good_friend_past_fct",
                 "friends_child_has_one_good_friend_last_six_months_fct",
                 "friends_yp_has_a_best_friend_fct")

test <- map(friends_var, function(x) tab_desc_tab(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])
```

```{r suppress = T}
map(friends_var, function(x) fig_desc_long(.data[[x]]))
```


```{r results='asis', suppress = T}
# cont vars
alspac_long %>%
  select(age, friends_score, friends_score_log) %>% 
  rename_all(~str_replace_all(., "_", ".")) %>% 
  group_by(age) %>% 
  summarise_all(list(mean = ~mean(., na.rm = T),
                sd = ~sd(., na.rm = T),
                miss = ~mean(is.na(.)) * 100)) %>% 
  na.omit() %>% 
  pivot_longer(!age,
                names_sep = "_",
                names_to = c("Variable", ".value")) %>%
  mutate(Variable = str_replace_all(Variable, "\\.", "_")) %>% 
  arrange(Variable, age) %>% 
  select(Variable, age, everything()) %>% 
  rename_all(~str_to_title(.)) %>% 
  knitr::kable(digits = 1)
```

```{r}
qplot(alspac_long$friends_score) + theme_bw()
qplot(alspac_long$friends_score_log) +theme_bw()

fig_cont_long(friends_score)


```






### School

```{r results='asis'}

school_vars <- c("school_talk_friends_fct", "school_likes_teacher_fct",
                 "school_left_out_fct", "school_everybodys_friend_fct", 
                 "school_classmates_not_all_their_friends_fct", 
                 "school_all_pupils_are_friends_fct", 
                 "school_some_classmates_mean_fct", 
                 "school_likes_schoolmates_fct", 
                 "school_frightened_by_schoolmates_fct", 
                 "school_feel_popular_fct", 
                 "school_get_on_with_classmates_fct", 
                 "school_schoolmates_accept_them_fct", 
                 "school_feel_lonely_fct", "school_people_trust_them_fct", 
                 "school_people_depend_on_them_fct")

test <- map(school_vars, function(x) tab_desc_tab(.data[[x]]))

for (i in seq_along(test)) print(test[[i]])
```

```{r suppress = T}
map(school_vars, function(x) fig_desc_long(.data[[x]]))
```






### Parents

```{r results='asis'}
alspac_long %>%
  select(age, parent_networks_mother_social_support_score, 
         parent_networks_mother_social_networks_score,
         parent_networks_partner_social_networks_score,
         parent_networks_partner_social_support_score) %>% 
  rename_all(~str_replace_all(., "_", ".")) %>% 
  group_by(age) %>% 
  summarise_all(list(mean = ~mean(., na.rm = T),
                     sd = ~sd(., na.rm = T),
                     miss = ~mean(is.na(.)) * 100)) %>% 
  pivot_longer(!age,
               names_sep = "_",
               names_to = c("Variable", ".value")) %>%
  mutate(Variable = str_replace_all(Variable, "\\.", "_")) %>% 
  arrange(Variable, age) %>% 
  select(Variable, age, everything()) %>% 
  rename_all(~str_to_title(.)) %>% 
  filter(!is.na(Sd)) %>% 
  knitr::kable(digits = 1)
```

```{r}
qplot(alspac_long$parent_networks_mother_social_support_score) +
  theme_bw()
qplot(alspac_long$parent_networks_mother_social_networks_score) +
  theme_bw()
qplot(alspac_long$parent_networks_partner_social_networks_score) +
  theme_bw()
qplot(alspac_long$parent_networks_partner_social_support_score) +
  theme_bw()

fig_cont_long(parent_networks_mother_social_support_score)
fig_cont_long(parent_networks_mother_social_networks_score)
fig_cont_long(parent_networks_partner_social_networks_score)
fig_cont_long(parent_networks_partner_social_support_score)
```


### Peer problem and social cohesion

```{r results='asis'}
alspac_long %>%
  select(age, peer_problem_score, social_cohesion, social_discord) %>% 
  rename_all(~str_replace_all(., "_", ".")) %>% 
  group_by(age) %>% 
  summarise_all(list(mean = ~mean(., na.rm = T),
                     sd = ~sd(., na.rm = T),
                     miss = ~mean(is.na(.)) * 100)) %>% 
  pivot_longer(!age,
               names_sep = "_",
               names_to = c("Variable", ".value")) %>%
  mutate(Variable = str_replace_all(Variable, "\\.", "_")) %>% 
  arrange(Variable, age) %>% 
  select(Variable, age, everything()) %>% 
  rename_all(~str_to_title(.)) %>% 
  filter(!is.na(Sd)) %>% 
  knitr::kable(digits = 1)
```


```{r}
qplot(alspac_long$peer_problem_score) +
  theme_bw() # skewed
qplot(alspac_long$social_cohesion) +
  theme_bw() 

fig_cont_long(peer_problem_score)
fig_cont_long(social_cohesion)
```







### Romantic

```{r results='asis', suppress = T}
romantic_var <- c("romantic_in_relationship_or_out_with_someone_fct",
                  "romantic_in_relationship_fct",
                  "romantic_emotionally_close_fct")

test <- map(romantic_var, function(x) tab_desc_tab(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])
```

```{r suppress = T}
map(romantic_var, function(x) fig_desc_long(.data[[x]]))
```


### Religion

```{r results='asis', supress = T}
religion_var <- c("religion_support_mother_fct",
                  "religion_support_partner_fct")

test <- map(religion_var, function(x) tab_desc_tab(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])
```

```{r suppress = T}
map(religion_var, function(x) fig_desc_long(.data[[x]]))
```


#### Neighbour



```{r}
qplot(alspac_long$neighbour_stress_score) +
  theme_bw()
fig_cont_long(neighbour_stress_score)
```

### Family

```{r results='asis'}



family_cat_var <- c("family_how_close_yp_feels_to_their_siblings",
                    "family_frequency_child_visits_relatives",
                    "family_child_sees_grandparents",
                    "family_child_gets_on_well_with_rest_of_family",
                    "family_child_sees_his_or_her_grandparents")

alspac_long %>%
  select(age, starts_with("family")) %>%
  select(-family_cat_var, -ends_with("fct")) %>% 
  rename_all(~str_replace_all(., "_", ".")) %>% 
  group_by(age) %>% 
  summarise_all(list(mean = ~mean(., na.rm = T),
                     sd = ~sd(., na.rm = T),
                     miss = ~mean(is.na(.)) * 100)) %>% 
  pivot_longer(!age,
               names_sep = "_",
               names_to = c("Variable", ".value")) %>%
  mutate(Variable = str_replace_all(Variable, "\\.", "_")) %>% 
  arrange(Variable, age) %>% 
  select(Variable, age, everything()) %>% 
  rename_all(~str_to_title(.)) %>% 
  filter(!is.na(Sd)) %>% 
  knitr::kable()



```

```{r results='asis', supress = T}
family_cat_var <- str_c(family_cat_var, "_fct")

test <- map(family_cat_var, function(x) tab_desc_tab(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])
```


```{r suppress = T}
map(family_cat_var, function(x) fig_desc_long(.data[[x]]))
```


### Childcare


```{r supress = T, results='asis'}
childcare_var <- alspac_long %>% 
  select(starts_with("childcare")) %>% 
  select(ends_with("fct")) %>% 
  names()

test <- map(childcare_var, function(x) tab_desc_tab(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])
```

```{r supress = T}
test <- map(childcare_var, function(x) fig_desc_long(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])
```


### Relationship


```{r supress = T, results='asis'}
relationship_var <- alspac_long %>% 
  select(starts_with("relationship")) %>% 
  select(ends_with("fct")) %>% 
  names()

test <- map(relationship_var, function(x) tab_desc_tab(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])

```

```{r supress = T}
test <- map(relationship_var, function(x) fig_desc_long(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])
```

```{r results='asis'}
alspac_long %>%
  select(age, relationship_to_parents_partner_positive_relationship_score, 
         relationship_to_parents_partner_negative_relationship_score,
         relationship_to_parents_partner_interaction_score,
         relationship_to_parents_mother_interaction_score) %>% 
  rename_all(~str_replace_all(., "_", ".")) %>% 
  group_by(age) %>% 
  summarise_all(list(mean = ~mean(., na.rm = T),
                     sd = ~sd(., na.rm = T),
                     miss = ~mean(is.na(.)) * 100)) %>% 
  na.omit() %>% 
  pivot_longer(!age,
               names_sep = "_",
               names_to = c("Variable", ".value")) %>%
  mutate(Variable = str_replace_all(Variable, "\\.", "_")) %>% 
  arrange(Variable, age) %>% 
  select(Variable, age, everything()) %>% 
  rename_all(~str_to_title(.)) %>% 
  knitr::kable(digits = 1)
```

```{r}
qplot(alspac_long$relationship_to_parents_partner_positive_relationship_score) +
  theme_bw()
qplot(alspac_long$relationship_to_parents_partner_negative_relationship_score) +
  theme_bw()
qplot(alspac_long$relationship_to_parents_partner_interaction_score) +
  theme_bw()
qplot(alspac_long$relationship_to_parents_mother_interaction_score) +
  theme_bw()
```


### Social


```{r supress = T, results='asis'}
social_var <- alspac_long %>% 
  select(starts_with("social")) %>% 
  select(ends_with("fct")) %>% 
  names()

test <- map(social_var, function(x) tab_desc_tab(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])

```

```{r supress = T}
test <- map(social_var, function(x) fig_desc_long(.data[[x]]))
for (i in seq_along(test)) print(test[[i]])
```


### Siblings

```{r results='asis'}
tab_desc_tab(sibling_interaction_child_argues_with_siblings_fct)
tab_desc_tab(sibling_interaction_child_argues_with_older_child_fct)
```

```{r}
fig_desc_long(sibling_interaction_child_argues_with_siblings_fct)
fig_desc_long(sibling_interaction_child_argues_with_older_child_fct)
```

```{r results='asis'}
alspac_long %>%
  select(age, sibling_interaction_sibling_interaction_score) %>% 
  rename_all(~str_replace_all(., "_", ".")) %>% 
  group_by(age) %>% 
  summarise_all(list(mean = ~mean(., na.rm = T),
                     sd = ~sd(., na.rm = T),
                     miss = ~mean(is.na(.)) * 100)) %>% 
  na.omit() %>% 
  rename_all(~str_to_title(.)) %>% 
  knitr::kable(digits = 1, 
               caption = "Sibling interaction score")
```



```{r}
qplot(alspac_long$sibling_interaction_sibling_interaction_score) +
  theme_bw()
```


## Control variables

```{r results='asis'}
count(alspac_long, white) %>% 
  mutate(Prop = n/sum(n),
         white = factor(white, labels = c("No", "Yes")),
         white = fct_explicit_na(white, na_level  = "Missing")) %>% 
  rename_all(~str_to_title(.)) %>% 
  knitr::kable(digits = 2, caption = "Ethnicity")
           

count(alspac_long, special_edu_needs) %>% 
  mutate(Prop = n/sum(n),
         special_edu_needs = factor(special_edu_needs, labels = c("No", "Yes")),
         special_edu_needs = fct_explicit_na(special_edu_needs, 
                                             na_level  = "Missing")) %>% 
  rename_all(~str_to_title(.)) %>% 
  knitr::kable(digits = 2, caption = "Special educational needs")


count(alspac_long, mom_edu_fct) %>% 
  mutate(Prop = n/sum(n),
         mom_edu_fct = fct_explicit_na(mom_edu_fct, 
                                       na_level  = "Missing")) %>% 
  rename_all(~str_to_title(.)) %>% 
  knitr::kable(digits = 2, caption = "Mother's education")

count(alspac_long, mom_partner) %>% 
  mutate(Prop = n/sum(n),
         mom_partner = factor(mom_partner, labels = c("No", "Yes")),
         mom_partner = fct_explicit_na(mom_partner, 
                                             na_level  = "Missing")) %>% 
  rename_all(~str_to_title(.)) %>% 
  knitr::kable(digits = 2, caption = "Mother has partner")

count(alspac_long, mom_work) %>% 
  mutate(Prop = n/sum(n),
         mom_work = factor(mom_work, labels = c("No", "Yes")),
         mom_work = fct_explicit_na(mom_work, 
                                             na_level  = "Missing")) %>% 
  rename_all(~str_to_title(.)) %>% 
  knitr::kable(digits = 2, caption = "Mother works")


```


```{r}



count(alspac_long, imd2000) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "IMD score 2000")

count(alspac_long, imd2004) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "IMD score 2004")

count(alspac_long, imd2007) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "IMD score 2007")

count(alspac_long, imd2010) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "IMD score 2010")

count(alspac_long, imd2015) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "IMD score 2015")


```


```{r}
# health describe
health_vars <- str_c("health_", c(8, 9, 11, 12, 14), "_fct")

alspac_long %>% 
  filter(age == -1) %>% 
  count(health_8_fct) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "Health score age 8")


alspac_long %>% 
  filter(age == -1) %>% 
  count(health_9_fct) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "Health score age 9")


alspac_long %>% 
  filter(age == -1) %>% 
  count(health_11_fct) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "Health score age 11")


alspac_long %>% 
  filter(age == -1) %>% 
  count(health_12_fct) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "Health score age 12")


alspac_long %>% 
  filter(age == -1) %>% 
  count(health_14_fct) %>% 
  setNames(c("Score", "N")) %>% 
  mutate(Prop = N/sum(N)) %>% 
  knitr::kable(digits = 2, caption = "Health score age 14")


```





# Bivariate statistics 

  

## Bullying

```{r}
bully_var <- c("bullying_victim_fct", 
               "bullying_often_bullied_fct",
               "bullying_frequency_bullied_fct")

alspac_long2 <- alspac_long %>% 
  select(uniqid, age, anxiety, depression, bullying_victim_fct,
         depression_smfq) %>%
  group_by(uniqid) %>% 
  mutate(bully_victim_8 = ifelse(age == 8 & bullying_victim_fct == "Yes",
                                 1, 0),
         bully_victim_8 = max(bully_victim_8, na.rm = T),
         bully_victim_10 = ifelse(age == 10 & bullying_victim_fct == "Yes",
                                 1, 0),
         bully_victim_10 = max(bully_victim_10, na.rm = T)) %>% 
  ungroup()



trends_bully_8 <- alspac_long2 %>% 
  group_by(age, bully_victim_8) %>% 
  summarise(Anxiety_prop = mean(anxiety, na.rm = T),
            Depression_prop = mean(depression, na.rm = T),
            Depress_smfq_prop = mean(depression_smfq, na.rm = T)) %>% 
#  filter(!is.nan(Anxiety_prop)) %>% 
  rename(bully = bully_victim_8) %>% 
  mutate(bully_age = 8) %>% 
  ungroup()


trends_bully_10 <- alspac_long2 %>% 
  group_by(age, bully_victim_10) %>% 
  summarise(Anxiety_prop = mean(anxiety, na.rm = T),
            Depression_prop = mean(depression, na.rm = T),
            Depress_smfq_prop = mean(depression_smfq, na.rm = T)) %>% 
#  filter(!is.nan(Anxiety_prop)) %>% 
  rename(bully = bully_victim_10) %>% 
  mutate(bully_age = 10) %>% 
  ungroup()

trends_bully <- rbind(trends_bully_8, trends_bully_10)


trends_bully %>% 
  gather(value = value, key = var, -age, -bully, -bully_age) %>%
  na.omit() %>% 
  mutate(var = str_remove(var, "_prop"),
         age = age,
         bully = ifelse(bully == 1, "Yes", "No"),
         bully = as.factor(bully) %>% fct_rev(),
         bully_age = ifelse(bully_age == 8,
                            "Bullied at age 8",
                            "Bullied at age 10"),
         bully_age = as.factor(bully_age) %>% fct_rev()) %>% 
  ggplot(aes(age, value, color = bully, 
             group = bully)) + 
  geom_line() +
  geom_point() +
  facet_grid(var~bully_age, scales = "free") +
  theme_bw() +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = unique(trends_bully$age)) +
  labs(x = "Age",
       y = "Percentage of cases",
       color = "Bullied") 


```

